---
title: "Train Combined BANTER Model"
author: "Selene Fregosi"
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: margin = 1cm
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    # code_folding: hide
    # keep_md: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      # cache = FALSE, 
                      # warning = FALSE, 
                      message = FALSE
)
```

## Purpose

To train a BANTER model with towed array, longline, and glider data. 

## Set up

### Load libraries
```{r libraries}
library(banter)
library(rfPermute)
library(ggplot2)
library(beepr)
```

### Set paths and run type
```{r set-paths}
# path to save models
# path_models <- file.path('T:/glider_MHI_analysis', 'classification', 'models')
# path_models <- file.path('Z:/LLHARP/analysis/banterModels')
path_models <- file.path('T:/fregosi/cross_platform_banter_big_data/models')

path_big_data <- 'T:/fregosi/cross_platform_banter_big_data'

# Re-generate/train model (1) or load existing model to view results (0)
freshRun <- 1
```

## Load training data

Training data was created for each platform individually

Alternatively, if this is not a fresh run to train a new model, load an existing model and just look at the results. 

```{r load-data}

if (freshRun == 1){
  tf_ta <- file.path('R:/McCullough/PAMpal/Banter_TowedArray_Data/', 
                     'banterAllsmall_HICEAS2017-CH5_Filtered-577.rdata')
  tf_ll <- file.path('T:/fregosi/cross_platform_banter_big_data', 
                     'banterDetsTrain_LLHARP_2024-05-27.rda')
  tf_gl <- file.path('T:/fregosi/cross_platform_banter_big_data', 
                     'banterDetsTrain_glider_2024-05-28.rda')
  
  #initialize events df
  allEvents <-NULL
  allDetectors <-NULL
  load(tf_ta)
  
  # change species codes
  idx577 <- which(banterAllsmall$events$species == '577')
  banterAllsmall$events$species[idx577] <- 'UO'
  idx33 <- which(banterAllsmall$events$species == '33')
  banterAllsmall$events$species[idx33] <- 'Pc'
  
  # assemble events
  allEvents <- banterAllsmall$events
  # assemble detectors
  allDetectors <-banterAllsmall$detectors
  
  # add longline
  load(tf_ll)
  allEvents <- rbind(allEvents, banterDetsTrain$events)
  allDetectors <- c(allDetectors, banterDetsTrain$detectors)
  
  # add glider
  load(tf_gl)
  allEvents <- rbind(allEvents, banterDetsTrain$events)
  allDetectors <- c(allDetectors, banterDetsTrain$detectors)
  
  # collapse the data.frames
  allDetectors <- tapply(allDetectors, names(allDetectors), dplyr::bind_rows)
  
  banterDetsTrain <- list(events = allEvents, detectors = allDetectors)
  save(banterDetsTrain, file = file.path(path_big_data, 
                                       paste0('banterDetsTrain_combined_', 
                                              Sys.Date(), '.rda')))
  
} else if (freshRun == 0){
  # modelFile <- file.path(path_models, 'LLHARP_BANTER_2class_2023-05-31.rda')
  modelFile <- file.path(path_models, 'bantMDL_combined_2024-05-28.rda')
  load(modelFile)
  cat('Loaded existing model:', modelFile, '\n')
}

```

## Clean data

If the training data contains species other than just `Pc` or `UO`, they need to be cleaned up. Replace all non-Pc entries with `UO`. 

Also, remove any 'Click_Detector_0' detections. 

```{r clean}

if (freshRun == 1){
  # uSp <- unique(allEvents$species)
    uSp <- unique(banterDetsTrain$events$species)
  uSp <- uSp[-which(uSp == 'UO' | uSp == 'Pc')]
  
  for (u in seq_along(uSp)){
    # uSpIdx <- which(allEvents$species == uSp[u])
    # allEvents$species[uSpIdx] <- 'UO'
        uSpIdx <- which(banterDetsTrain$events$species == uSp[u])
    banterDetsTrain$events$species[uSpIdx] <- 'UO'
  }
  
  # unique(allEvents$species)
  
  # Check for click_detector_0 in training dataset
  # allDetectors$Click_Detector_0 <- NULL
  banterDetsTrain$detectors$Click_Detector_0 <- NULL
}

```

## Build the model
```{r build}

if (freshRun == 1){
  # build the model
  bant.mdl <- initBanterModel(banterDetsTrain$events)
  bant.mdl <- addBanterDetector(
    bant.mdl,
    data =  banterDetsTrain$detectors, 
    ntree = 10000,
    sampsize = 5,
    importance = TRUE,
    num.cores = 1
  )
  bant.mdl <- runBanterModel(bant.mdl, ntree = 10000, sampsize = 5)
  
  # save model
  save(bant.mdl, file = file.path(path_models, paste0('bantMDL_combined_',
                                                      Sys.Date(), '_ss5_nt10k.rda')))
  
  # get random forest data and save
  event.rf <- getBanterModel(bant.mdl, "event")
  save(event.rf, file = file.path(path_models, paste0('eventRF_combined_',
                                                      Sys.Date(), '_ss5_nt10k.rda')))
} else if (freshRun == 0){
  event.rf <- getBanterModel(bant.mdl, "event")
  cat('No model trained. Showing results from existing model:', modelFile, '\n')
}

# beepr::beep()

```

## View results

### Overall results (event model)
```{r results-overall}

summary(bant.mdl)
plotVotes(event.rf)
plotProximity(event.rf)

```

### Detector results (detector models)
```{r results-detectors}
detList <- names(bant.mdl@detectors)

for (d in seq_along(detList)){
  cat(detList[d], '\n')
  plotTrace(getBanterModel(bant.mdl, model = detList[d]))
  #plotInbag(getBanterModel(bant.mdl, model = d))
  # print(table(getBanterModel(bant.mdl, model = d)$oob.times))
}

```




