---
title: "Train Glider BANTER Model"
author: "Selene Fregosi"
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: margin = 1cm
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    # code_folding: hide
    # keep_md: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      # cache = FALSE, 
                      # warning = FALSE, 
                      message = FALSE
)
```

## Purpose

To train a BANTER model on glider-collected acoustic data. It is trained on data from two missions by a single glider/recorder system: sg639_MHI_Apr2022 and sg649_MHI_Apr2023. Events were identified via manual scanning of LTSAs in Triton and Pamguard was run across all recordings using the standard BANTER suite of click, whistle and moan, and cepstrum detectors. AcousticStudy objects were created for each mission using `PAMpal` and applied the same filtering as the longline and towed array datasets. These AcousticStudies were combined using `gl3_combine_acoustic_studies.R` to create the final trainig dataset. 

## Set up

### Load libraries
```{r libraries}
library(banter)
library(rfPermute)
library(ggplot2)
library(beepr)
```

### Set paths and run type
```{r set-paths}
# path to save models
# path_models <- file.path('T:/glider_MHI_analysis', 'classification', 'models')
# path_models <- file.path('Z:/LLHARP/analysis/banterModels')
path_models <- file.path('T:/fregosi/cross_platform_banter_big_data/models')

# Re-generate/train model (1) or load existing model to view results (0)
freshRun <- 1
```

## Load training data

Training data was created with `gl3_combine_acoustic_studies.R`. 

Alternatively, if this is not a fresh run to train a new model, load an existing model and just look at the results. 

```{r load-data}

if (freshRun == 1){
  trainingFile <- file.path('T:/fregosi/cross_platform_banter_big_data', 
                            'banterDetsTrain_glider_2024-05-28.rda')
  load(trainingFile)
  # load to-be-predicted file
  # TBPFile <-'Z:/LLHARP/analysis/banter2024/simonis_data/AllEvents_TBP_Pc_UO_Pm2023-05-30.rda'
  # load(TBPFile)
} else if (freshRun == 0){
  # modelFile <- file.path(path_models, 'LLHARP_BANTER_2class_2023-05-31.rda')
  modelFile <- file.path(path_models, 'bantMDL_glider_2024-05-28.rda')
  load(modelFile)
  cat('Loaded existing model:', modelFile, '\n')
}

```

## Clean data

If the training data contains species other than just `Pc` or `UO`, they need to be cleaned up. Replace all non-Pc entries with `UO`. 

Also, remove any 'Click_Detector_0' detections. 

```{r clean}

if (freshRun == 1){
  uSp <- unique(banterDetsTrain$events$species)
  uSp <- uSp[-which(uSp == 'UO' | uSp == 'Pc')]
  
  for (u in seq_along(uSp)){
    uSpIdx <- which(banterDetsTrain$events$species == uSp[u])
    banterDetsTrain$events$species[uSpIdx] <- 'UO'
  }
  
  # unique(banterDetsTrain$events$species)
  
  # Check for click_detector_0 in training dataset
  banterDetsTrain$detectors$Click_Detector_0 <- NULL
}

```

## Build the model
```{r build}

if (freshRun == 1){
  # build the model
  bant.mdl <- initBanterModel(banterDetsTrain$events)
  bant.mdl <- addBanterDetector(
    bant.mdl,
    data =  banterDetsTrain$detectors, 
    ntree = 12000,
    sampsize = 5,
    importance = TRUE,
    num.cores = 1
  )
  bant.mdl <- runBanterModel(bant.mdl, ntree = 10000, sampsize = 5)
  
  # save model
  save(bant.mdl, file = file.path(path_models, paste0('bantMDL_glider_',
                                                      Sys.Date(), '.rda')))
  
  # get random forest data and save
  event.rf <- getBanterModel(bant.mdl, "event")
  save(event.rf, file = file.path(path_models, paste0('eventRF_glider_',
                                                      Sys.Date(), '.rda')))
} else if (freshRun == 0){
  event.rf <- getBanterModel(bant.mdl, "event")
  cat('No model trained. Showing results from existing model:', modelFile, '\n')
}

# beepr::beep()

```

## View results

### Overall results (event model)
```{r results-overall}

summary(bant.mdl)
plotVotes(event.rf)
plotProximity(event.rf)

```

### Detector results (detector models)
```{r results-detectors}
detList <- names(bant.mdl@detectors)

for (d in seq_along(detList)){
  cat(detList[d], '\n')
  plotTrace(getBanterModel(bant.mdl, model = detList[d]))
  #plotInbag(getBanterModel(bant.mdl, model = d))
  # print(table(getBanterModel(bant.mdl, model = d)$oob.times))
}

```




